Why is it that statistical tests for residuals are not widely used? An explanation using visual inference.

### 2023-01-22 paper_comparison/paper_comparison/paper_comparison.Rmd
fix plot in the different fitted value distribution section

## Effect of other factors

@@

(calc the diff in power and plot)

(randomly subset uniform and check the change of the power)

Although the distribution of the regressor is classically not restricted by the assumptions of linear regression, it will have an impact on the distribution of the fitted values and the visual effect of the residual plot. In our experiments, four distributions of the regressor are used to accommodate this variation.

Figure \ref{fig:different-x-dist-poly-power} and Figure \ref{fig:different-x-dist-heter-power} illustrate the change of power of visual tests and conventional tests for different distribution of $X_{raw}$ used in the non-linearity model and the heteroskedasticity model. We focus on $log(E) > 2$ as the visual patterns become visible.

For the non-linearity model, we do not observe significant difference of power of visual test between distributions, except for the discrete uniform distribution. It could be due to the fact that humans have difficulties in recognising visual patterns from discreteness, which makes the shape disconnected and incomplete (ref?). In terms of conventional tests, the discrete uniform  distribution is instead the one with the greatest power, followed by the lognormal distribution, uniform distribution and the normal distribution.

@@

In regards of the type of the simulated model, the power of visual tests under different distributions of regressor are always lower than the corresponding conventional tests, consistent with the result from the previous sections.

```{r eval = FALSE}
vi_survey %>%
  filter(!null_lineup) %>%
  filter(!attention_check) %>%
  filter(type == "polynomial") %>%
  filter(x_dist == "even_discrete") %>%
  filter(log(effect_size) > 4) %>%
  filter(p_value > 0.05) %>%
  .$unique_lineup_id %>%
  unique()
```

```{r eval = FALSE}
VI_MODEL$plot_lineup(vi_lineup$poly_64$data, 
                     remove_grid_line = TRUE, 
                     theme = theme_light(),
                     remove_axis = TRUE)
```

@@

dat_different_dist <- poly_conv_sim %>% 
  mutate(log_effect_size = log(effect_size)) %>%
  select(x_dist, log_effect_size, RESET4_p_value) %>%
  rename(RESET_p_value = RESET4_p_value) %>%
  pivot_longer(RESET_p_value) %>%
  mutate(x_dist = ifelse(x_dist == "even_discrete", 
                                       "Discrete", 
                                       stringr::str_to_title(x_dist))) %>%
  mutate(name = gsub("_p_value", " test", name)) %>%
  mutate(reject = value < 0.05) %>%
  mutate(conv_or_not = "Conventional") %>%
  bind_rows(vi_survey %>%
  group_by(unique_lineup_id) %>%
  summarise(across(everything(), first)) %>%
  filter(type == "polynomial") %>%
  filter(!null_lineup) %>%
  filter(!attention_check) %>%
  mutate(reject = as.numeric(p_value <= 0.05)) %>%
  mutate(log_effect_size = log(effect_size)) %>%
  mutate(x_dist = ifelse(x_dist == "even_discrete", 
                         "Discrete", 
                         stringr::str_to_title(x_dist))) %>%
  mutate(conv_or_not = "Visual")) %>%
  mutate(x_dist = factor(x_dist, 
      levels = c("Uniform", "Normal", 
                 "Lognormal", "Discrete")))

ggplot() +
  geom_smooth(data=filter(dat_different_dist, 
                          x_dist == "Uniform"),
              aes(log_effect_size, 
                  reject), 
                  col = "grey", 
              method = "glm", 
              method.args = list(family = binomial), 
              se = FALSE, size=3) +
  geom_smooth(data=dat_different_dist,
              aes(log_effect_size, 
                  reject, 
                  col = x_dist), 
              method = "glm", 
              method.args = list(family = binomial), 
              se = FALSE, size=1) +
  theme_light(base_size = 5) +
  scale_color_manual(values = rcartocolor::carto_pal(4, "Safe")[c(4,1,3,2)]) +
  scale_size_manual(values = c(1, 2, 3, 4)) +
  xlab(expression(log[e] (Effect_size))) +
  ylab("Power") +
  labs(col = "", linetype = "Test") +
  facet_wrap(~conv_or_not) +
  theme(legend.position="bottom")
```

```{r different-x-dist-heter-power, fig.height = 3, fig.width = 5, fig.cap = "Power of conventional tests and visual tests on lineups produced by the heteroskedasticity model under different distributions of regressor."}

dat_different_dist <- heter_conv_sim %>% 
  mutate(log_effect_size = log(effect_size)) %>%
  select(x_dist, log_effect_size, BP_p_value) %>%
  pivot_longer(BP_p_value) %>%
  mutate(x_dist = ifelse(x_dist == "even_discrete", 
                                       "Discrete", 
                                       stringr::str_to_title(x_dist))) %>%
  mutate(name = gsub("_p_value", " test", name)) %>%
  mutate(reject = value < 0.05) %>%
  mutate(conv_or_not = "Conventional") %>%
  bind_rows(vi_survey %>%
  group_by(unique_lineup_id) %>%
  summarise(across(everything(), first)) %>%
  filter(type == "heteroskedasticity") %>%
  filter(!null_lineup) %>%
  filter(!attention_check) %>%
  mutate(reject = as.numeric(p_value <= 0.05)) %>%
  mutate(log_effect_size = log(effect_size)) %>%
  mutate(x_dist = ifelse(x_dist == "even_discrete", 
                         "Discrete", 
                         stringr::str_to_title(x_dist))) %>%
  mutate(conv_or_not = "Visual"))

dat_different_dist %>%
  ggplot() +
  geom_smooth(aes(log_effect_size, 
                  reject, 
                  col = x_dist), 
              method = "glm", 
              method.args = list(family = binomial), 
              se = FALSE) +
  theme_light(base_size = 5) +
  scale_color_manual(values = rcartocolor::carto_pal(4, "Safe")[c(1,3,2,4)]) +
  scale_size_manual(values = c(1, 2, 3)) +
  xlab(expression(log[e] (Effect_size))) +
  ylab("Power") +
  labs(col = "Distribution of regressor", linetype = "Test") +
  facet_wrap(~conv_or_not)
```

### Shape

>>>>

## Effect of other factors used in the simulated model

@@

Although the distribution of the regressor is classically not restricted by the assumptions of linear regression, it will have an impact on the fitted value distribution and the visual effect of the residual plot. In our experiments, four distributions of fitted values are considered to accommodate this variation.

Figure \ref{fig:different-x-dist-poly-power} and Figure \ref{fig:different-x-dist-heter-power} illustrate the change of power of visual tests and conventional tests for different fitted distributions used in the non-linearity model and the heteroskedasticity model. We focus on $log(E) > 2$ where the visual patterns start to become recognisable.

For the non-linearity model, we do not observe significant difference in power of visual test between distributions, except for the discrete uniform distribution. It is possibility due to the fact that humans have difficulties in recognising visual patterns from discreteness, which makes the shape disconnected and incomplete (ref?). In terms of conventional tests, the discrete uniform distribution is instead the one with the greatest power, followed by the lognormal distribution, the uniform distribution and the normal distribution.

@@

In regards of the type of the simulated model, the power of visual tests under different distributions of regressor are always lower than the corresponding conventional tests, which is consistent with the result from the previous sections.

@@

dat_different_dist <- poly_conv_sim %>% 
  mutate(log_effect_size = log(effect_size)) %>%
  select(x_dist, log_effect_size, RESET4_p_value) %>%
  rename(RESET_p_value = RESET4_p_value) %>%
  pivot_longer(RESET_p_value) %>%
  mutate(x_dist = ifelse(x_dist == "even_discrete", 
                                       "Discrete", 
                                       stringr::str_to_title(x_dist))) %>%
  mutate(name = gsub("_p_value", " test", name)) %>%
  mutate(reject = value < 0.05) %>%
  mutate(conv_or_not = "Conventional") %>%
  bind_rows(vi_survey %>%
  group_by(unique_lineup_id) %>%
  summarise(across(everything(), first)) %>%
  filter(type == "polynomial") %>%
  filter(!null_lineup) %>%
  filter(!attention_check) %>%
  mutate(reject = as.numeric(p_value <= 0.05)) %>%
  mutate(log_effect_size = log(effect_size)) %>%
  mutate(x_dist = ifelse(x_dist == "even_discrete", 
                         "Discrete", 
                         stringr::str_to_title(x_dist))) %>%
  mutate(conv_or_not = "Visual")) %>%
  mutate(x_dist = factor(x_dist, 
      levels = c("Uniform", "Normal", 
                 "Lognormal", "Discrete"))) %>%
  select(x_dist, log_effect_size, reject, conv_or_not)

dat_different_dist %>%
  ggplot() +
  geom_smooth(data = ~filter(.x, x_dist == "Uniform"),
              aes(log_effect_size, 
                  reject), 
                  col = "grey", 
              method = "glm", 
              method.args = list(family = binomial), 
              se = FALSE, size = 3) +
  geom_smooth(aes(log_effect_size, 
                  reject, 
                  col = x_dist), 
              method = "glm", 
              method.args = list(family = binomial), 
              se = FALSE, 
              size = 1) +
  theme_light(base_size = 5) +
  scale_color_manual(values = rcartocolor::carto_pal(4, "Safe")[c(4,1,3,2)]) +
  scale_size_manual(values = c(1, 2, 3, 4)) +
  xlab(expression(log[e] (Effect_size))) +
  ylab("Power") +
  labs(col = "", linetype = "Test") +
  facet_wrap(~conv_or_not) +
  theme(legend.position="bottom")
```

```{r different-x-dist-heter-power, fig.height = 3, fig.width = 7, fig.cap = "Power of conventional tests and visual tests on lineups produced by the heteroskedasticity model under different distributions of regressor."}

dat_different_dist <- heter_conv_sim %>% 
  mutate(log_effect_size = log(effect_size)) %>%
  select(x_dist, log_effect_size, BP_p_value) %>%
  pivot_longer(BP_p_value) %>%
  mutate(x_dist = ifelse(x_dist == "even_discrete", 
                                       "Discrete", 
                                       stringr::str_to_title(x_dist))) %>%
  mutate(name = gsub("_p_value", " test", name)) %>%
  mutate(reject = value < 0.05) %>%
  mutate(conv_or_not = "Conventional") %>%
  bind_rows(vi_survey %>%
  group_by(unique_lineup_id) %>%
  summarise(across(everything(), first)) %>%
  filter(type == "heteroskedasticity") %>%
  filter(!null_lineup) %>%
  filter(!attention_check) %>%
  mutate(reject = as.numeric(p_value <= 0.05)) %>%
  mutate(log_effect_size = log(effect_size)) %>%
  mutate(x_dist = ifelse(x_dist == "even_discrete", 
                         "Discrete", 
                         stringr::str_to_title(x_dist))) %>%
  mutate(conv_or_not = "Visual")) %>%
  mutate(x_dist = factor(x_dist, 
      levels = c("Uniform", "Normal", 
                 "Lognormal", "Discrete"))) %>%
  select(x_dist, log_effect_size, reject, conv_or_not)

dat_different_dist %>%
  ggplot() +
  geom_smooth(data = ~filter(.x, x_dist == "Uniform"),
              aes(log_effect_size, 
                  reject), 
                  col = "grey", 
              method = "glm", 
              method.args = list(family = binomial), 
              se = FALSE, size = 3) +
  geom_smooth(aes(log_effect_size, 
                  reject, 
                  col = x_dist), 
              method = "glm", 
              method.args = list(family = binomial), 
              se = FALSE, 
              size = 1) +
  theme_light(base_size = 5) +
  scale_color_manual(values = rcartocolor::carto_pal(4, "Safe")[c(4,1,3,2)]) +
  scale_size_manual(values = c(1, 2, 3, 4)) +
  xlab(expression(log[e] (Effect_size))) +
  ylab("Power") +
  labs(col = "", linetype = "Test") +
  facet_wrap(~conv_or_not) +
  theme(legend.position="bottom")
```

### Shape of non-linearity

### Shape of heteroskedasticity

### 2023-01-22 paper_comparison/paper_comparison/paper_comparison.tex
fix plot in the different fitted value distribution section

\hypertarget{effect-of-other-factors}{%
\subsection{Effect of other factors}\label{effect-of-other-factors}}

@@

(calc the diff in power and plot)

(randomly subset uniform and check the change of the power)

Although the distribution of the regressor is classically not restricted
by the assumptions of linear regression, it will have an impact on the
distribution of the fitted values and the visual effect of the residual
plot. In our experiments, four distributions of the regressor are used
to accommodate this variation.

Figure \ref{fig:different-x-dist-poly-power} and Figure
\ref{fig:different-x-dist-heter-power} illustrate the change of power of
visual tests and conventional tests for different distribution of
\(X_{raw}\) used in the non-linearity model and the heteroskedasticity
model. We focus on \(log(E) > 2\) as the visual patterns become visible.

For the non-linearity model, we do not observe significant difference of
power of visual test between distributions, except for the discrete
uniform distribution. It could be due to the fact that humans have
difficulties in recognising visual patterns from discreteness, which
makes the shape disconnected and incomplete (ref?). In terms of
conventional tests, the discrete uniform distribution is instead the one
with the greatest power, followed by the lognormal distribution, uniform
distribution and the normal distribution.

@@

In regards of the type of the simulated model, the power of visual tests
under different distributions of regressor are always lower than the
corresponding conventional tests, consistent with the result from the
previous sections.

@@

\hypertarget{shape}{%
\subsubsection{Shape}\label{shape}}

>>>>

\hypertarget{effect-of-other-factors-used-in-the-simulated-model}{%
\subsection{Effect of other factors used in the simulated
model}\label{effect-of-other-factors-used-in-the-simulated-model}}

@@

Although the distribution of the regressor is classically not restricted
by the assumptions of linear regression, it will have an impact on the
fitted value distribution and the visual effect of the residual plot. In
our experiments, four distributions of fitted values are considered to
accommodate this variation.

Figure \ref{fig:different-x-dist-poly-power} and Figure
\ref{fig:different-x-dist-heter-power} illustrate the change of power of
visual tests and conventional tests for different fitted distributions
used in the non-linearity model and the heteroskedasticity model. We
focus on \(log(E) > 2\) where the visual patterns start to become
recognisable.

For the non-linearity model, we do not observe significant difference in
power of visual test between distributions, except for the discrete
uniform distribution. It is possibility due to the fact that humans have
difficulties in recognising visual patterns from discreteness, which
makes the shape disconnected and incomplete (ref?). In terms of
conventional tests, the discrete uniform distribution is instead the one
with the greatest power, followed by the lognormal distribution, the
uniform distribution and the normal distribution.

@@

In regards of the type of the simulated model, the power of visual tests
under different distributions of regressor are always lower than the
corresponding conventional tests, which is consistent with the result
from the previous sections.

@@

\hypertarget{shape-of-non-linearity}{%
\subsubsection{Shape of non-linearity}\label{shape-of-non-linearity}}

\hypertarget{shape-of-heteroskedasticity}{%
\subsubsection{Shape of
heteroskedasticity}\label{shape-of-heteroskedasticity}}
